Children’s Vocabulary Development in American Sign Language

### 2023-04-19 outputs/paper/paper.qmd
added model to quarto

In this paper, we investigate the patterns and trends that occur during the early developmental years of childhood in terms of vocabulary in American Sign Language. The estimand that we are trying to estimate is the relationship between children\'s developmental ages and their production of vocabulary words in American Sign Language. We aimed to visualize the milestones for children\'s vocabulary development in sign language to compare it to the timeline of vocabulary development for speech-language. This paper uses data obtained from the open database, Wordbank, which exhibits information on the development of children\'s vocabulary in various languages. The dataset, \"Full Child-by-Word\", available on this database, provides information on the demographic details of each child when provided as well as the variables presenting their vocabulary assessment.

@@

@@

```{r}
#| label: fig-childbyactionwordprod
#| fig-cap: Child's Amount of Produced ASL Action Words
#| echo: false
#| warning: false

# 3 ## each child's amount of produced action words ##
produced_action_words <-
  cleaned_vocab_data |>
    filter(production_ability == "YES", item_category == "action_words")

produced_action_words |>
  ggplot(mapping = aes(x = child_id)) +
  geom_bar(fill = "blue") +
  labs(
    x = "Child ID",
    y = "Number of Action Words Produced",
    title = "Children's Ability to Produce ASL Action Words"
  )

@@

```{r}
#| label: fig-childbytimewordprod
#| fig-cap: Child's Amount of Produced ASL Time Words
#| echo: false
#| warning: false

# 4 ## each child's amount of produced time words ##
produced_time_words <-
  cleaned_vocab_data |>
    filter(production_ability == "YES", item_category == "time_words")

produced_time_words |>
  ggplot(mapping = aes(x = child_id)) +
  geom_bar(fill = "blue") +
  labs(
    x = "Child ID",
    y = "Number of Time Words Produced",
    title = "Children's Ability to Produce ASL Time Words"
  )

```{r}
#| label: fig-childbyquestionwordprod
#| fig-cap: Child's Amount of Produced ASL Question Words
#| echo: false
#| warning: false

# 5 ## each child's amount of produced question words ##
produced_question_words <-
  cleaned_vocab_data |>
  filter(production_ability == "YES", item_category == "question_words")

produced_question_words |>
  ggplot(mapping = aes(x = child_id)) +
  geom_bar(fill = "blue") +
  labs(
    x = "Child ID",
    y = "Number of Time Question Produced",
    title = "Children's Ability to Produce ASL Question Words"
  )

```{r}
#| label: fig-childbydescriptivewordprod
#| fig-cap: Child's Amount of Produced ASL Descriptive Words
#| echo: false
#| warning: false

# 6 ## each child's amount of produced descriptive words ##
produced_descriptive_words <-
  cleaned_vocab_data |>
  filter(production_ability == "YES", item_category == "descriptive_words")

produced_descriptive_words |>
  ggplot(mapping = aes(x = child_id)) +
  geom_bar(fill = "blue") +
  labs(
    x = "Child ID",
    y = "Number of Time Descriptive Produced",
    title = "Children's Ability to Produce ASL Descriptive Words"
  )

@@

@@

>>>>

In this paper, we investigate the patterns and trends that occur during the early developmental years of childhood in terms of vocabulary in American Sign Language. The estimand that we are trying to estimate is the relationship between children's developmental ages and their production of vocabulary words in American Sign Language. We aimed to visualize the milestones for children's vocabulary development in sign language to compare it to the timeline of vocabulary development for speech-language. This paper uses data obtained from the open database, Wordbank, which exhibits information on the development of children's vocabulary in various languages. The dataset, "Full Child-by-Word", available on this database, provides information on the demographic details of each child when provided as well as the variables presenting their vocabulary assessment.

@@

```{r}
#| echo: false

## summary statistics table
summary_stats_vocab_data <-
  cleaned_vocab_data |>
  mutate(
    vocabulary_item_definition = NULL,
    item_category = NULL,
    overall_assessment = NULL
  ) |>
  rename(
    "Production Ability" = production_ability,
    "Comprehension Ability" = comprehension_ability
  ) |>
  datasummary_skim(type = "categorical")

# present summary statistics table
summary_stats_vocab_data
```

@@

# READ MODEL INTO QUARTO DOCUMENT
prod_vocab_model <- readRDS(here::here("outputs/models/sim_prod_vocab_model.rds"))

@@

@@

```{=tex}
\begin{equation}
Y_{i} = \beta_{0}+\beta_{1}Age_{i} (\#eq:1)
\end{equation}
```
In Model \@ref(eq:1):

-   $Y$ is ...
-   $\beta_{0}$ is ...
-   $\beta_{1}$ is ...

@@

```{r}
#| label: fig-prodwordsmodel
#| fig-cap: Linear Model of Produced Words by Children Ages
#| echo: false
#| warning: false

prod_vocab_model |>
  ggplot(aes(x = age_in_months, y = prod_words)) +
  geom_point(alpha = 0.5) +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "black",
    linetype = "dashed",
    formula = "y ~ x"
  ) +
  labs(
    x = "Age (Months)",
    y = "Number of ASL Words Produced"
  ) +
  theme_classic()
```

```{r}
#| echo: false

modelsummary(prod_vocab_model)
```
